Short-time simulation of collective spin Hamiltonians with arbitrary decoherence

### 2018-10-13 soc_squeezing/cumulant_expansions.tex
made text in the main body more general, i.e. apply for any basis for collective spin operators

\title{Solving Collective Spin Hamiltonians with Decoherence}

@@

Past work has shown how to realize spin-squeezing Hamiltonians with a
variety of ultracold atomic and trapped ion systems.  While simulating
the dynamics and squeezing behavior induced by these Hamiltonians is
straightforward in the context of a closed quantum system, here we
develop a method to simulate dynamics in the presence of decoherence.
In particular, we will consider dynamics induced by the one-axis
twisting (OAT), two-axis twisting (TAT), and transverse field (TVF)
Hamiltonians of the form
\begin{align}
  H_{\t{OAT}} = \chi S_\z^2,
  &&
  H_{\t{TAT}}^{\z,\y} = \chi \p{S_\z^2 - S_\y^2},
  &&
  H_{\t{TVF}} = \chi S_\z^2 + \Omega S_\x.
\end{align}
In order to determine the squeezing behavior of these Hamiltonians, we
will study the dynamics of Heisenburg operators $\O$, whose evolution
under a Hamiltonian $H$ is given by
\begin{align}
  \f{d}{dt} \O = i\sp{H,\O}_- + \sum_\J \gamma_\J \D\p{\J;\O},
  &&
  \D\p{\J;\O}
  \equiv \sum_{J\in\J}\p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+},
  \label{eq:EOM}
\end{align}
where $\J$ is a set of jump operators with a corresponding decoherence
rate $\gamma_\J$ and $\sp{A,B}_\pm\equiv AB\pm BA$.  Our measure of
spin squeezing is
\begin{align}
  \xi^2 \equiv \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $N$ is the total number of spins, $\v S$ is the collective spin
vector, and the minimization is performed over all unit vectors
$\uv n$ in the plane orthogonal to the mean spin vector $\bk{\v S}$.
This squeezing parameter is entirely determined by components
$\bk{S_\alpha}$ of the mean spin vector in addition to collective
spin-spin correlators of the form $\bk{S_\alpha S_\beta}$; our task is
therefore to compute these expectation values in all scenarios of
interest.

@@

The case of the one-axis twisting is simplest to consider, as we can
write the Hamiltonian in the form
\begin{align}
  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_\z^{(j)} \sigma_\z^{(k)}
  = \f12 \chi \sum_{j<k} \sigma_\z^{(j)} \sigma_\z^{(k)} + \f14 N \chi,
\end{align}
which admits an exact solution previously worked out by Foss-Feig et
al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
expectation values with respect to the initial state $\ket{\X}$
satisfying $S_\x\ket{\X}=S\ket{\X}$, in terms of $S\equiv N/2$ and
$\mu,\nu\in\set{+1,-1}$ we find that
\begin{align}
  \bk{S_+}
  &= \sum_j\bk{\sigma_+^{(j)}}
  = S e^{-\Gamma t} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
  \bk{S_\mu S_\z}
  &= \f12\sum_j\bk{\sigma_\mu^{(j)}\sigma_\z^{(j)}}
  + \f12\sum_{j\ne k} \bk{\sigma_\mu^{(j)} \sigma_\z^{(k)}} \\
  &= -\f{\mu}{2}\bk{S_\mu} + S \p{S-\f12} e^{-\Gamma t}
  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
  \bk{S_\mu S_\nu}
  &= \sum_j \bk{\sigma_\mu^{(j)} \sigma_\nu^{(j)}}
  + \sum_{j\ne k} \bk{\sigma_\mu^{(j)} \sigma_\nu^{(k)}} \\
  &= \delta_{\mu,-\nu} \p{S + \mu\bk{S_\z}}
  + S \p{S-\f12} e^{-2\Gamma t}
  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
\end{align}
where
\begin{align}
  \Phi\p{X,t}
  &\equiv e^{-\lambda t} \sp{\cos\p{t\sqrt{s_X^2-r}}
    + \lambda t~\t{sinc}\p{t\sqrt{s_X^2-r}}},
  \\
  \Psi\p{X,t}
  &\equiv e^{-\lambda t} \p{is_X-\gamma}t~
  \t{sinc}\p{t\sqrt{s_X^2-r}},
\end{align}
for
\begin{align}
  \gamma \equiv \f12 \p{\gamma_+ - \gamma_-},
  &&
  \lambda \equiv \f12 \p{\gamma_+ + \gamma_-},
  &&
  r \equiv \gamma_+ \gamma_-,
  &&
  \Gamma \equiv \gamma_\z + \lambda,
  &&
  s_X \equiv X + i\gamma,
\end{align}
defined respectively in terms of spontaneous uncorrelated spin
dephasing, excitation, and decay rates $\gamma_\z$, $\gamma_+$,
$\gamma_-$.  The last ingredients we need to compute spin squeezing at
any time are $\bk{S_\z}$ and $\bk{S_\z^2}$.  As these operators
commute with the OAT Hamiltonian, their evolution is governed entirely
by decoherence (see Appendix \ref{sec:decay_single}), which means
\begin{align}
  \f{d}{dt} S_\z
  &= S\p{\gamma_+-\gamma_-} - \p{\gamma_++\gamma_-} S_\z,
  \\
  \f{d}{dt}\p{S_\z^2}
  &= S\p{\gamma_++\gamma_-} + 2\p{\gamma_+-\gamma_-}\p{S-\f12} S_\z
  - 2 \p{\gamma_++\gamma_-} S_\z^2.
\end{align}
The initial conditions $\bk{S_\z}_{t=0}=0$ and $\bk{S_\z^2}_{t=0}=S/2$
then imply
\begin{align}
  \bk{S_\z}
  = S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
  \p{1-e^{-\p{\gamma_+ + \gamma_-} t}},
  &&
  \bk{S_\z^2} = S \sp{\f12 + \p{S-\f12} \f{\bk{S_\z}^2}{S^2}}.
  \label{eq:Sz_OAT}
\end{align}
The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
\eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
$\xi^2$ defined in \eqref{eq:squeezing} at any time throughout
evolution of the initial state $\ket{\X}:S_\x\ket{\X}=S\ket{\X}$ under
$H_{\t{OAT}}$.

\section{Truncated operator dynamics}

Unlike the one-axis twisting Hamiltonian, the two-axis twisting and
transverse-field Hamiltonians do not admit exact analytical
solutions. We will therefore solve these Hamiltonians approximately
using alternate methods.  The basic strategy will be to find equations
of motion for operators of interest, and at some point truncate
high-order terms with negligible contribution to spin correlators of
interest.  To benchmark this method, we test it against exact one-axis
twisting results in the presence of decoherence, collective
Hamiltonian dynamics in the absence of decoherence, and exact
simulations of few-spin systems.

All collective spin Hamiltonians can be written in the form
\begin{align}
  H = \sum_{\ell,m,n} h_{\ell mn} S_+^\ell S_\z^m S_-^n,
  \label{eq:H_general}
\end{align}
where Hermiticity requires that $h_{\ell mn}=h_{nm\ell}^*$.  The
Hamiltonian for one-axis twisting with a transverse field, for
example, is
\begin{align}
  H_{\t{TVF}}
  = \chi S_\z^2 + \Omega S_\x
  = \chi S_\z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-,
\end{align}
while the two-axis twisting Hamiltonians about $\p{\z,\y}$ and
$\p{\y,\x}$ are
\begin{align}
  H_{\t{TAT}}^{\z,\y}
  \propto S_\z^2 - S_\y^2
  = S_\z^2 + \f14 S_+^2 + \f14 S_-^2
  - \f12 S_+ S_- + \f12 S_\z,
\end{align}
and
\begin{align}
  H_{\t{TAT}}^{\y,\x}
  \propto S_\y^2 - S_\x^2
  = -\f12 S_+^2 - \f12 S_-^2.
\end{align}
The evolution of a general operator $S_+^\ell S_\z^m S_-^n$ under a
Hamiltonian of the form in \eqref{eq:H_general} is given by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n}
  = i \sum_{p,q,r} h_{pqr}
  \sp{S_+^p S_\z^q S_-^r, S_+^\ell S_\z^m S_-^n}_-
  + \sum_\J \gamma_\J \D\p{\J; S_+^\ell S_\z^m S_-^n}
  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r,
  \label{eq:time_derivative}
\end{align}
where commutator and decoherence terms necessary for computing the
matrix elements $T^{\ell m n}_{pqr}$ of the time derivative operator
$d/dt$ are worked out Appendices \ref{sec:comm_general},
\ref{sec:decoherence_single}, and \ref{sec:decoherence_collective}.

The time derivative operator $d/dt$ will generally couple spin
operators $S_+^\ell S_\z^m S_-^n$ to spin operators
$S_+^p S_\z^q S_-^r$ with higher ``weight'', i.e.~with
$p+q+r>\ell+m+n$, which signifies the growth of many-body
correlations.  In practice, keeping track of this operator growth will
eventually require more computational resources than are available, so
we must truncate it at some point by taking
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n}
  = \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r
  \to \sum_{w\p{p,q,r}<W} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r
  \label{eq:time_derivative_truncation}
\end{align}
for some cutoff $W$ and weight function $w$, e.g.~$w\p{p,q,r}=p+q+r$.
The truncation in \eqref{eq:time_derivative_truncation} will close the
system of differential equations defined by
\eqref{eq:time_derivative}, and allow us to solve it using standard
numerical methods.  The initial conditions for this system of
differential equations, i.e.~expectation values of operators with
respect to states that are simple to prepare, are provided in Appendix
\ref{sec:initial_conditions}.

The truncation strategy in \eqref{eq:time_derivative_truncation} has a
few limitations:
\begin{enumerate*}[label=(\roman*)]
\item simulating a system of differential equations for a large number
  of operators can be very time-consuming,
\item the weight function $w$ may need to be chosen carefully, as the
  optimal weight function is generally system-dependent,
\item simulation results can only be trusted up to the time at which
  the initial value of operators $S_+^p S_\z^r S_-^q$ with weight
  $w\p{p,q,r}\ge W$ have a non-negligible contribution to operators of
  interest.\label{pt:limitation}
\end{enumerate*}
While the first two of these limitations might in principle be avoided
by alternate means of tracking operators, the last limitation
necessarily applies in some form to any method tracking only a subset
of {\it all} operators.  Embracing the unavoidable, we therefore try
to devise an alternate truncation strategy which is built around
\ref{pt:limitation}.

We can expand operators in a Taylor series about $t=0$ to write
\begin{align}
  S_+^\ell S_\z^m S_-^n
  = \sum_{k\ge0} \f{t^k}{k!}
  \left.\f{d^k}{dt^k}\right|_{t=0} \p{S_+^\ell S_\z^m S_-^n},
\end{align}
which implies
\begin{align}
  \bk{S_+^\ell S_\z^m S_-^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_\z^q S_-^r}_{t=0},
  \label{eq:time_series}
\end{align}
where the matrix elements of the $k$-th time derivative are given by
\begin{align}
  T^{\ell m n}_{pqr;0}=\delta_{\ell p}\delta_{mq}\delta_{nr},
  &&
  T^{\ell m n}_{pqr;1}\equiv T^{\ell m n}_{pqr},
\end{align}
\begin{align}
  T^{\ell mn}_{pqr;k>1}
  \equiv \sum_{\v\lambda,\v\mu,\v\nu}
  T^{\ell mn}_{\lambda_1 \mu_1 \nu_1}
  T^{\lambda_1 \mu_1 \nu_1}_{\lambda_2 \mu_2 \nu_2}
  T^{\lambda_2 \mu_2 \nu_2}_{\lambda_3 \mu_3 \nu_3}
  \cdots T^{\lambda_{k-1} \mu_{k-1} \nu_{k-1}}_{pqr}.
\end{align}
Instead of explicitly choosing a weight function that determines which
operators to keep track of, we can now truncate the series in
\eqref{eq:time_series} by taking
\begin{align}
  \bk{S_+^\ell S_\z^m S_-^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_\z^q S_-^r}_{t=0}
  \to \sum_{k<M} \f{t^k}{k!}
  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_\z^q S_-^r}_{t=0}.
\end{align}
Examining the nonzero entries of $T^{\ell mn}_{pqr;k}$ for
$k=0,1,\cdots,M-1$ now {\it tells} us which operators we need to keep
track of in order to compute the expectation value
$\bk{S_+^\ell S_\z^m S_-^n}$.  Furthermore, after computing all
relevant matrix elements $T^{\ell mn}_{pqr;k}$ and expectation values
$\bk{S_+^p S_\z^q S_-^r}_{t=0}$, computing
$\bk{S_+^\ell S_\z^m S_-^n}$ takes only a short time independent of
system size.  Finally, the relation
$\p{S_+^\ell S_\z^m S_-^n}^\dag=S_+^n S_\z^m S_-^\ell$ can be
exploited to cut both the number of matrix elements
$T^{\ell mn}_{pqr}$ and expectation values
$\bk{S_+^p S_\z^r S_-^q}_{t=0}$ that we need to compute roughly in
half.

\section{Decoherence in the two-axis twisting protocols}

In the case of squeezing through spin-orbit coupling with the Sr-87
optical lattice clock, the primary form of decoherence is through
decay of individual spins, i.e.~with jump operators $\sigma_-^{(j)}$.
The two axis twisting Hamiltonians $H_{\t{TAT}}^{\z,\y}$ and
$H_{\t{TAT}}^{\z,\y}$, however, are actually realized in a rotating
frame, which implies that we must transform these jump operators
appropriately.  Specifically, the TAT protocol begins with a
time-dependent transverse-field Hamiltonian
\begin{align}
  H = \chi S_\z^2 + \Omega\p{t} S_\x,
  &&
  \Omega\p{t} = \beta\omega \cos\p{\omega t},
\end{align}
where $\beta$ is the modulation index of the drive, and the drive
frequency $\omega\gg N\chi$, with $N$ the total number of spins.
Moving into the rotating frame of $\Omega\p{t}S_\x$ subtracts this
term from the Hamiltonian, and transforms operators $\O$ as
\begin{align}
  \O \to U\p{t}^\dag \O U\p{t},
  &&
  U\p{t} \equiv \exp\sp{-i\int_0^t\d\tau~\Omega\p{\tau}S_\x}
  = \exp\sp{-i\beta\sin\p{\omega t}S_\x},
\end{align}
In particular, the operators $\tilde S_\pm\equiv-S_\z\pm iS_\y$
transform simply as
\begin{align}
  \tilde S_\pm
  \to U^\dag \tilde S_\pm U
  = e^{\pm i\beta\sin\p{\omega t}} \tilde S_\pm.
\end{align}
For any operator $\O$ and drive frequency $\omega\gg\norm{\O}$, where
$\norm{\O}$ denotes the operator norm of $\O$, we can generally make a
secular approximation to say
\begin{align}
  e^{\pm im\beta\sin\p{\omega t}} \O
  = \sum_{n=-\infty}^\infty J_n\p{\pm m\beta} e^{in\omega t} \O
  \approx J_0\p{\pm m\beta} \O
  = J_0\p{m\beta} \O,
\end{align}
where $J_n$ is the $n$-th order Bessel function of the first kind.
Expanding $S_\z^2=\f14\p{\tilde S_+ + \tilde S_-}^2$, one can thus
work out that the effective Hamiltonian in the rotating frame of the
drive is
\begin{align}
  H_{\t{eff}}
  \approx \f{\chi}{2} \p{\sp{1+J_0\p{2\beta}} S_\z^2
    + \sp{1-J_0\p{2\beta}} S_\y^2}.
\end{align}
Driving with a modulation index $\beta$ for which
$J_0\p{2\beta}=\pm1/3$ then gives us the effective two-axis twisting
Hamiltonians
\begin{align}
  H_{\t{eff}}^{(+)}
  = \f{\chi}{3} \p{2 S_\z^2 + S_\y^2}
  \simeq \f{\chi}{3} \p{S_\z^2 - S_\x^2},
  &&
  H_{\t{eff}}^{(-)}
  = \f{\chi}{3} \p{S_\z^2 + 2 S_\y^2}
  \simeq \f{\chi}{3} \p{S_y^2 - S_\x^2},
\end{align}
where $\simeq$ denotes equality up to overall factors of
$\v S^2=S_\z^2+S_\x^2+S_\y^2$ which are irrelevant in the absence of
coherent coupling between states with different net spin.  In a
similar spirit, one can work out that single-spin operators transverse
to the $\x$-axis transform as
\begin{align}
  \tilde\sigma_\pm^{(j)}
  \equiv \f12\p{-\sigma_\z^{(j)}\pm i\sigma_\y^{(j)}}
  \to U^\dag \tilde\sigma_\pm^{(j)} U
  = e^{\pm i\beta\sin\p{\omega t}} \tilde\sigma_\pm^{(j)}
  \approx J_0\p{\beta} \tilde\sigma_\pm^{(j)},
\end{align}
which implies
\begin{align}
  \sigma_\z^{(j)} \to J_0\p{\beta} \sigma_\z^{(j)},
  &&
  \sigma_\pm^{(j)}
  \to \f12\sp{1 \pm J_0\p{\beta}} \sigma_+^{(j)}
  + \f12\sp{1 \mp J_0\p{\beta}} \sigma_-^{(j)}.
\end{align}
The TAT protocol about $\p{\z,\y}$ requires
\begin{align}
  J_0\p{2\beta} = 1/3
  \implies \beta \approx 0.906
  \implies J_0\p{\beta} \approx 0.805,
\end{align}
while the TAT protocol about $\p{\y,\x}$ requires
\begin{align}
  J_0\p{2\beta} = -1/3
  \implies
  \begin{cases}
    \beta \approx 1.63
    \implies J_0\p{\beta} \approx 0.440, \\
    \beta \approx 2.22
    \implies J_0\p{\beta} \approx 0.099.
  \end{cases}
\end{align}

@@

ordering on products of spin operators, which will allow us to only

@@

mutually dependent intermediate variables!  We therefore devote the
rest of this section to simplifying our result for the commutator

@@

In this section we work out the effects of decoherence via collective
jump operators on ordered products of collective spin operators.  For
shorthand, we define
\begin{align}
  \D\p{\Gamma;\O}
  \equiv \D\p{\set{\Gamma};\O}
  = \Gamma^\dag \O \Gamma - \f12\sp{\Gamma^\dag \Gamma, \O}_+,
\end{align}
where $\Gamma$ is a collective spin operator with weight one i.e.~a
linear combination of $S_\z$, $S_\mu$, and $S_\bmu$.

@@

\section{Changing operator bases}

>>>>

\title{Short-time simulation of collective spin Hamiltonians with
  arbitrary decoherence}

@@

Previous work has shown how to realize spin-squeezing Hamiltonians
with a variety of ultracold atomic and trapped ion systems.  While
simulating the dynamics of initially Gaussian states evolving under
these Hamiltonians is straightforward in the context of a closed
quantum system, here we develop a method to simulate short-time
dynamics in the presence of decoherence.  In particular, we will
consider dynamics induced by the one-axis twisting (OAT), two-axis
twisting (TAT), and transverse field (TVF) Hamiltonians of the form
\begin{align}
  H_{\t{OAT}} = \chi S_\z^2,
  &&
  H_{\t{TAT}}^{\z,\y} = \chi \p{S_\z^2 - S_\y^2},
  &&
  H_{\t{TVF}} = \chi S_\z^2 + \Omega S_\x.
\end{align}
In order to determine the squeezing behavior of these Hamiltonians, we
will study the dynamics of Heisenburg operators $\O$, whose evolution
under a Hamiltonian $H$ is given by
\begin{align}
  \f{d}{dt} \O = i\sp{H,\O}_- + \sum_\J \gamma_\J \D\p{\J;\O},
  &&
  \D\p{\J;\O}
  \equiv \sum_{J\in\J}\p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+},
  \label{eq:EOM}
\end{align}
where $\J$ is a set of jump operators with a corresponding decoherence
rate $\gamma_\J$ and $\sp{X,Y}_\pm\equiv XY\pm YX$.  Our measure of
spin squeezing is
\begin{align}
  \xi^2 \equiv \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $N$ is the total number of spins, $\v S$ is the collective spin
vector, and the minimization is performed over all unit vectors
$\uv n$ in the plane orthogonal to the mean spin vector $\bk{\v S}$.
This squeezing parameter is entirely determined by components
$\bk{S_\alpha}$ of the mean spin vector in addition to collective
spin-spin correlators of the form $\bk{S_\alpha S_\beta}$; our task is
therefore to compute these expectation values in all scenarios of
interest.

@@

The simplest case is that of one-axis twisting, as we can write the
Hamiltonian in the form
\begin{align}
  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_\z^{(j)} \sigma_\z^{(k)}
  = \f12 \chi \sum_{j<k} \sigma_\z^{(j)} \sigma_\z^{(k)} + \f14 N \chi,
\end{align}
which admits an exact solution previously worked out by Foss-Feig et
al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
expectation values with respect to the initial state $\ket{\X}$
satisfying $S_\x\ket{\X}=S\ket{\X}$, in terms of $S\equiv N/2$ and
$\mu,\nu\in\set{+1,-1}$ we find that
\begin{align}
  \bk{S_+}
  &= \sum_j\bk{\sigma_+^{(j)}}
  = S e^{-\Gamma t} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
  \bk{S_\mu S_\z}
  &= \f12\sum_j\bk{\sigma_\mu^{(j)}\sigma_\z^{(j)}}
  + \f12\sum_{j\ne k} \bk{\sigma_\mu^{(j)} \sigma_\z^{(k)}} \\
  &= -\f{\mu}{2}\bk{S_\mu} + S \p{S-\f12} e^{-\Gamma t}
  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
  \bk{S_\mu S_\nu}
  &= \sum_j \bk{\sigma_\mu^{(j)} \sigma_\nu^{(j)}}
  + \sum_{j\ne k} \bk{\sigma_\mu^{(j)} \sigma_\nu^{(k)}} \\
  &= \delta_{\mu,-\nu} \p{S + \mu\bk{S_\z}}
  + S \p{S-\f12} e^{-2\Gamma t}
  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
\end{align}
where
\begin{align}
  \Phi\p{X,t}
  &\equiv e^{-\lambda t} \sp{\cos\p{t\sqrt{s_X^2-r}}
    + \lambda t~\t{sinc}\p{t\sqrt{s_X^2-r}}},
  \\
  \Psi\p{X,t}
  &\equiv e^{-\lambda t} \p{is_X-\gamma}t~
  \t{sinc}\p{t\sqrt{s_X^2-r}},
\end{align}
for
\begin{align}
  \gamma \equiv \f12 \p{\gamma_+ - \gamma_-},
  &&
  \lambda \equiv \f12 \p{\gamma_+ + \gamma_-},
  &&
  r \equiv \gamma_+ \gamma_-,
  &&
  \Gamma \equiv \gamma_\z + \lambda,
  &&
  s_X \equiv X + i\gamma,
\end{align}
defined respectively in terms of uncorrelated spin dephasing,
excitation, and decay rates $\gamma_\z$, $\gamma_+$, $\gamma_-$.  The
last ingredients we need to compute spin squeezing at any time are
$\bk{S_\z}$ and $\bk{S_\z^2}$.  As these operators commute with the
OAT Hamiltonian, their evolution is governed entirely by decoherence
(see Appendix \ref{sec:decay_single}), which means
\begin{align}
  \f{d}{dt} S_\z
  &= S\p{\gamma_+-\gamma_-} - \p{\gamma_++\gamma_-} S_\z,
  \\
  \f{d}{dt}\p{S_\z^2}
  &= S\p{\gamma_++\gamma_-} + 2\p{\gamma_+-\gamma_-}\p{S-\f12} S_\z
  - 2 \p{\gamma_++\gamma_-} S_\z^2.
\end{align}
The initial conditions $\bk{S_\z}_{t=0}=0$ and $\bk{S_\z^2}_{t=0}=S/2$
then imply
\begin{align}
  \bk{S_\z}
  = S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
  \p{1-e^{-\p{\gamma_+ + \gamma_-} t}},
  &&
  \bk{S_\z^2} = S \sp{\f12 + \p{S-\f12} \f{\bk{S_\z}^2}{S^2}}.
  \label{eq:Sz_OAT}
\end{align}
The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
\eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
$\xi^2$ defined in \eqref{eq:squeezing} at any time throughout
evolution of the initial state $\ket{\X}$ under $H_{\t{OAT}}$.

\section{Short-time evolution via truncated equations of motion}

Unlike the one-axis twisting Hamiltonian, the two-axis twisting and
transverse-field Hamiltonians do not admit exact analytical solutions.
We will therefore simulate dynamics under these Hamiltonians
numerically using alternate methods.  The basic strategy will be to
first find equations of motion for all operators of interest using
some complete basis for collective-spin operators.  These equations of
motion will generally yield a system of equations involving a number
of operators that grows too fast with system size for exact
simulation.  We will therefore look for ways to truncate the equations
of motion to arrive at a reasonably-sized expressions that are
practical to compute.  To benchmark this method, we test it against
collective Hamiltonian dynamics in the absence of decoherence, exact
one-axis twisting results in the presence of decoherence, and exact
simulations of small systems.

Ultimately, we are interested in the dynamics of collective spin
operators.  Choosing a basis $\set{\S_{\v m}}$ for these operators,
e.g.~$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
$\v m\equiv\p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$, we can expand all
collective-spin Hamiltonians in the form
\begin{align}
  H = \sum_{\v m} h_{\v m} \S_{\v m},
  \label{eq:H_general}
\end{align}
where if $\S_{\v m}=S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ then Hermiticity
requires that $h_{m_+,m_\z,m_-}=h_{m_-,m_\z,m_+}^*$.  The one-axis
twisting Hamiltonian with a transverse field, for example, is
\begin{align}
  H_{\t{TVF}}
  = \chi S_\z^2 + \Omega S_\x
  = \chi S_\z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-,
\end{align}
while the two-axis twisting Hamiltonians about $\p{z,y}$ and $\p{y,x}$
are
\begin{align}
  H_{\t{TAT}}^{\z,\y}
  \propto S_\z^2 - S_\y^2
  = S_\z^2 + \f14 S_+^2 + \f14 S_-^2
  - \f12 S_+ S_- + \f12 S_\z,
\end{align}
and
\begin{align}
  H_{\t{TAT}}^{\y,\x}
  \propto S_\y^2 - S_\x^2
  = -\f12 S_+^2 - \f12 S_-^2.
\end{align}
The evolution of a general operator $\S_{\v m}$ under a Hamiltonian of
the form in \eqref{eq:H_general} is given by
\begin{align}
  \f{d}{dt} \S_{\v m}
  = i \sum_{\v n} h_{\v n}
  \sp{\S_{\v n}, \S_{\v m}}_-
  + \sum_\J \gamma_\J \D\p{\J; \S_{\v m}}
  \equiv \sum_{\v n} T_{\v m\v n} \S_{\v n},
  \label{eq:time_derivative}
\end{align}
where commutator and decoherence terms necessary for computing the
matrix elements $T_{\v m\v n}$ of the time derivative operator $d/dt$
are worked out Appendices \ref{sec:comm_general},
\ref{sec:decoherence_single}, and \ref{sec:decoherence_collective}.

The time derivative operator $d/dt$ will generally couple spin
operators $\S_{\v m}$ to spin operators $\S_{\v n}$ with higher
``weight'', i.e.~with $\abs{\v n}>\abs{\v m}$ where
$\abs{\v\ell}\equiv\sum_\lambda\ell_\lambda$.  The growth of operator
weight signifies the growth of many-body correlations.  In practice,
keeping track of this growth will eventually require more
computational resources than are available, so we truncate it by
taking
\begin{align}
  \f{d}{dt} \S_{\v m}
  = \sum_{\v n} T_{\v m\v n} \S_{\v n}
  \to \sum_{w\p{\v n}<W} T_{\v m\v n} \S_{\v n}
  \label{eq:time_derivative_truncation}
\end{align}
for some weight measure $w$, e.g.~$w\p{\v n}=\abs{\v n}$, and cutoff
$W$.  The truncation in \eqref{eq:time_derivative_truncation} will
close the system of differential equations defined by
\eqref{eq:time_derivative}, and allow us to solve it using standard
numerical methods.  The initial conditions for this system of
differential equations, i.e.~expectation values of operators with
respect to states that are simple to prepare, are provided in Appendix
\ref{sec:initial_conditions}.

The truncation strategy in \eqref{eq:time_derivative_truncation} has a
few limitations:
\begin{enumerate*}[label=(\roman*)]
\item simulating a system of differential equations for a large number
  of operators can be very time-consuming,
\item the weight measure $w$ may need to be chosen carefully, as the
  optimal measure is generally system-dependent,
\item simulation results can only be trusted up to the time at which
  the initial value of operators $\S_{\v m}$ with weight
  $w\p{\v m}\ge W$ have a non-negligible contribution to operators of
  interest.\label{pt:limitation}
\end{enumerate*}
The last limitation in particular {\it necessarily} applies in some
form to any method tracking only a subset of all relevant operators.
Embracing the unavoidable, we therefore try to devise an alternate
truncation strategy built around \ref{pt:limitation}.

We can expand Heisenberg operators in a Taylor series about $t=0$ to
write
\begin{align}
  \bk{\S_{\v m}}
  = \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0},
  \label{eq:time_series}
\end{align}
where the matrix elements of the $k$-th time derivative are given by
\begin{align}
  T_{\v m\v n;0} \equiv \delta_{\v m\v n},
  &&
  T_{\v m\v n;1} \equiv T_{\v m\v n},
  &&
  T_{\v m\v n;k>1}
  \equiv \sum_{\v p_1,\v p_2,\cdots\v p_{k-1}}
  T_{\v m\v p_1} T_{\v p_1\v p_2} T_{\v p_2\v p_3}
  \cdots T_{\v p_{k-1}\v n}.
\end{align}
Instead of explicitly choosing a weight function that determines which
operators to keep track of, we can now truncate the series in
\eqref{eq:time_series} by taking
\begin{align}
  \bk{\S_{\v m}}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0}
  \to \sum_{k=0}^M \f{t^k}{k!}
  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0}.
\end{align}
The nonzero entries of $T_{\v m\v n;k}$ for $k=0,1,\cdots,M$ now tell
us which operators we need to keep track of in order to compute the
expectation value $\bk{\S_{\v m}}$.  Furthermore, after computing all
relevant matrix elements $T_{\v m\v n;k}$ and expectation values
$\bk{\S_{\v n}}_{t=0}$, computing $\bk{\S_{\v m}}$ takes only a short
time independent of system size.  Finally, the relation
$\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$ can be
exploited to cut both the number of matrix elements $T_{\v m\v n}$ and
expectation values $\bk{\S_{\v m}}_{t=0}$ that we need to compute
explicitly roughly in half.

@@

ordering on products of spin operators, which will ensure that we only

@@

mutually dependent intermediate variables, each term of which
additionally contains a product of $p$ factors!  We therefore devote
the rest of this section to simplifying our result for the commutator

@@

In this section we work out the effects of decoherence via collective
jump operators on ordered products of collective spin operators.  For
shorthand, we define
\begin{align}
  \D\p{\Gamma;\O}
  \equiv \D\p{\set{\Gamma};\O}
  = \Gamma^\dag \O \Gamma - \f12\sp{\Gamma^\dag \Gamma, \O}_+,
\end{align}
where $\Gamma$ is a collective spin operator with weight one, i.e.~a
linear combination of $S_\z$, $S_\mu$, and $S_\bmu$.

@@

\section{Changing bases for collective-spin operators}

For a collective-spin system composed of many small spins, the
operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ keep track
of $\abs{\v m}$-body operators (for $\abs{\v m}\equiv\sum_\mu m_\mu$),
but also carry lots of ``baggage'' in the form of $k$-body operators
with $k<\abs{\v m}$.  To cut down on this overhead, we change our
basis for collective-spin operators from $\S_{\v m}$ to the
purely-$\abs{\v m}$-body operators
\begin{align}
  \tilde\S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
  &&
  P_{\v j} \equiv \prod_{\mu,a} \sigma_\mu^{\p{j^\mu_a}},
\end{align}
where $\v m \equiv \p{m_\z,m_+,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
number of $\sigma_\mu$ operators in each term of $\S_{\v m}$; $\v j$
is an ordered list of indices for the spins addressed by $P_{\v j}$;
and the sum over $\p{\v j;\v m}$ denotes a sum over all possible
$\v j$ for which all indices in $\v j$ are distinct and
$\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
spin addressed by a $\sigma_\mu$ operator in $P_{\v j}$.

We now expand a product of basis operators $\tilde\S_{\v m}$ and
